---
title: "Survival of C.elegans with three different media"
author: "Jonas Gehrlein"
date: "12 nov 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(survival) 
library(here)
library(survminer)
```
code for fitting Kaplan-Meier and log-rank test and for displaying 
survival curves for each type of media and mutant.

```{r}
data <- read.table(here('data','Worm_mutant.csv'), header = TRUE, sep = ";", dec = ",")
head(data)
```

here() should show that your position is in the folder 7.semester else select the active project to be 7.semester in the upper right corner. 

Or make a new .Rproj file in the folder 7.semester. There is a problem with the numbers of the worms so we change that with col.names()
```{r}
colnames(data)<-c("Time","status",'Replicate', 'media',
                  'mutant')
head(data)
```

Then we create a survival object with the survival package and make a kaplan-meier curve

First we see for each mutant
```{r}
d_RRF <- subset(data, mutant == 'RRF')
d_PMK <- subset(data,mutant == 'PMK')
d_DAF <- subset(data, mutant == 'DAF')
```
```{r}
d_RRF$Survobj <- with(d_RRF,Surv(d_RRF$Time, event = d_RRF$status))
km_RRF <- survfit(Survobj ~ media, data = d_RRF,conf.type = "log-log", error = "greenwood")
s_km_RRF <- summary(km_RRF)
ggsurvplot(km_RRF,data = d_RRF, conf.int = TRUE, 
           ggtheme = theme_bw(),risk.table = 0.25,
           palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media',break.x.by = 2,
           title = 'RRF-3') 
```

```{r}
df_fly_13 <- data.frame(c(0,s_km_RRF$time[1:6]),c(1,s_km_RRF$surv[1:6]),                        c(0,s_km_RRF$std.err[1:6]))
colnames(df_fly_13) <- c('Time','Surv','Std.error')
df_fly_20 <- data.frame(c(0,s_km_RRF$time[7:11])
                        ,c(1,s_km_RRF$surv[7:11]),    c(0,s_km_RRF$std.err[7:11]))
colnames(df_fly_20) <- c('Time','Surv','Std.error')
df_OP50 <- data.frame(c(0,s_km_RRF$time[12:17])
                      ,c(1,s_km_RRF$surv[12:17]),
                        c(0,s_km_RRF$std.err[12:17]))
colnames(df_OP50) <- c('Time','Surv','Std.error')

plot(df_fly_13$Time[2:7],df_fly_13$Surv[2:7], pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival heat stress for'~italic(C.elegans)),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
lines(df_fly_13$Time,df_fly_13$Surv)
arrows(df_fly_13$Time, df_fly_13$Surv-df_fly_13$Std.error, df_fly_13$Time, 
       df_fly_13$Surv+df_fly_13$Std.error, length=0.05, angle=90, code=3, col = 'black')
points(df_OP50$Time[2:9],df_OP50$Surv[2:9], pch = 1, cex = 1.2)
lines(df_OP50$Time,df_OP50$Surv, lty = 2)
arrows(df_OP50$Time, df_OP50$Surv-df_OP50$Std.error,
      df_OP50$Time, df_OP50$Surv+df_OP50$Std.error, 
      length=0.05, angle=90, code=3, col = 'black')
points(df_fly_20$Time[2:9],df_fly_20$Surv[2:9], pch = 6)
lines(df_fly_20$Time,df_fly_20$Surv,lty = 3)
arrows(df_fly_20$Time, df_fly_20$Surv-df_fly_20$Std.error, df_fly_20$Time, 
       df_fly_20$Surv+df_fly_20$Std.error, length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,6,1),lty = c(1,3,2), 
       legend = c(expression('Fly gut 13'*degree*C),
                  expression('Fly gut 20'*degree*C),
                  expression(italic(E.coli) ~ OP50)), bty = 'n')
```

```{r}
d_PMK$Survobj <- with(d_PMK,Surv(d_PMK$Time, event = d_PMK$status))
km_PMK <- survfit(Survobj ~ media, data = d_PMK,conf.type = "log-log", error = "greenwood")
s_km_PMK <- summary(km_PMK)
ggsurvplot(km_PMK,data = d_PMK, conf.int = TRUE, 
           ggtheme = theme_bw(),risk.table = 0.25, 
           palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media',break.x.by = 2,
           title = 'PMK-1') 
```

```{r}
df_fly_13 <- data.frame(c(0,s_km_PMK$time[1:5]),
                        c(1,s_km_PMK$surv[1:5]),
                        c(0,s_km_PMK$std.err[1:5]))
colnames(df_fly_13) <- c('Time','Surv','Std.error')
df_fly_20 <- data.frame(c(0,s_km_PMK$time[6:10]),
                        c(1,s_km_PMK$surv[6:10]),
                        c(0,s_km_PMK$std.err[6:10]))
colnames(df_fly_20) <- c('Time','Surv','Std.error')
df_OP50 <- data.frame(c(0,s_km_PMK$time[11:17]),
                      c(1,s_km_PMK$surv[11:17]),
                        c(0,s_km_PMK$std.err[11:17]))
colnames(df_OP50) <- c('Time','Surv','Std.error')

plot(df_fly_13$Time[2:9],df_fly_13$Surv[2:9], pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival heat stress for'~italic(C.elegans)),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
lines(df_fly_13$Time,df_fly_13$Surv)
arrows(df_fly_13$Time, df_fly_13$Surv-df_fly_13$Std.error, df_fly_13$Time, 
       df_fly_13$Surv+df_fly_13$Std.error, length=0.05, angle=90, code=3, col = 'black')
points(df_OP50$Time[2:9],df_OP50$Surv[2:9], pch = 1, cex = 1.2)
lines(df_OP50$Time,df_OP50$Surv, lty = 2)
arrows(df_OP50$Time, df_OP50$Surv-df_OP50$Std.error,
      df_OP50$Time, df_OP50$Surv+df_OP50$Std.error, 
      length=0.05, angle=90, code=3, col = 'black')
points(df_fly_20$Time[2:9],df_fly_20$Surv[2:9], pch = 6)
lines(df_fly_20$Time,df_fly_20$Surv,lty = 3)
arrows(df_fly_20$Time, df_fly_20$Surv-df_fly_20$Std.error, df_fly_20$Time, 
       df_fly_20$Surv+df_fly_20$Std.error, length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,6,1),lty = c(1,3,2), 
       legend = c(expression('Fly gut 13'*degree*C),
                  expression('Fly gut 20'*degree*C),
                  expression(italic(E.coli) ~ OP50)), bty = 'n')
```

```{r}
d_DAF$Survobj <- with(d_DAF,Surv(d_DAF$Time, event = d_DAF$status))
km_DAF <- survfit(Survobj ~ media, data = d_DAF,conf.type = "log-log", error = "greenwood")
s_km_DAF <- summary(km_DAF)
ggsurvplot(km_DAF,data = d_DAF, conf.int = TRUE, 
           ggtheme = theme_bw(),risk.table = 0.25,
           palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media',break.x.by = 2,
           title = 'DAF-16') 
```

```{r}
df_fly_13 <- data.frame(c(0,s_km_DAF$time[1:3]),
                        c(1,s_km_DAF$surv[1:3]),
                        c(0,s_km_DAF$std.err[1:3]))
colnames(df_fly_13) <- c('Time','Surv','Std.error')
df_fly_20 <- data.frame(c(0,s_km_DAF$time[4:8]),
                        c(1,s_km_DAF$surv[4:8]),
                        c(0,s_km_DAF$std.err[4:8]))
colnames(df_fly_20) <- c('Time','Surv','Std.error')
df_OP50 <- data.frame(c(0,s_km_DAF$time[9:15]),
                      c(1,s_km_DAF$surv[9:15]),
                        c(0,s_km_DAF$std.err[9:15]))
colnames(df_OP50) <- c('Time','Surv','Std.error')

plot(df_fly_13$Time[2:9],df_fly_13$Surv[2:9], pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival heat stress for'~italic(C.elegans)),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
lines(df_fly_13$Time,df_fly_13$Surv)
arrows(df_fly_13$Time, df_fly_13$Surv-df_fly_13$Std.error, df_fly_13$Time, 
       df_fly_13$Surv+df_fly_13$Std.error, length=0.05, angle=90, code=3, col = 'black')
points(df_OP50$Time[2:9],df_OP50$Surv[2:9], pch = 1, cex = 1.2)
lines(df_OP50$Time,df_OP50$Surv, lty = 2)
arrows(df_OP50$Time, df_OP50$Surv-df_OP50$Std.error,
      df_OP50$Time, df_OP50$Surv+df_OP50$Std.error, 
      length=0.05, angle=90, code=3, col = 'black')
points(df_fly_20$Time[2:9],df_fly_20$Surv[2:9], pch = 6)
lines(df_fly_20$Time,df_fly_20$Surv,lty = 3)
arrows(df_fly_20$Time, df_fly_20$Surv-df_fly_20$Std.error, df_fly_20$Time, 
       df_fly_20$Surv+df_fly_20$Std.error, length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,6,1),lty = c(1,3,2), 
       legend = c(expression('Fly gut 13'*degree*C),
                  expression('Fly gut 20'*degree*C),
                  expression(italic(E.coli) ~ OP50)), bty = 'n')
```
Then we compare by medium
```{r}
data_OP50 <- subset(data, media == 'OP50')
data_FG13 <- subset(data, media == 'FG13')
data_FG20 <- subset(data, media == 'FG20')
```

```{r}
data_OP50$Survobj <- with(data_OP50,Surv(data_OP50$Time, event = data_OP50$status))
km_OP50 <- survfit(Survobj ~ mutant, data = data_OP50,conf.type = "log-log", error = "greenwood")
s_km_OP50 <- summary(km_OP50)
ggsurvplot(km_OP50,data = data_OP50, conf.int = TRUE, 
           ggtheme = theme_bw(),risk.table = 0.25, 
           palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media',break.x.by = 2,
           title = 'OP50') 
```


```{r}
data_FG13$Survobj <- with(data_FG13,Surv(data_FG13$Time, event = data_FG13$status))
km_FG13 <- survfit(Survobj ~ mutant, data = data_FG13,conf.type = "log-log", error = "greenwood")
s_km_FG13 <- summary(km_FG13)
ggsurvplot(km_FG13,data = data_FG13, conf.int = TRUE, 
           ggtheme = theme_bw(),risk.table = 0.25, 
           palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media',break.x.by = 2,
           title = 'FG13') 
```

```{r}
data_FG20$Survobj <- with(data_FG20,Surv(data_FG20$Time, event = data_FG20$status))
km_FG20 <- survfit(Survobj ~ mutant, data = data_FG20,conf.type = "log-log", error = "greenwood")
s_km_FG20 <- summary(km_FG20)
ggsurvplot(km_FG20,data = data_FG20, conf.int = TRUE, 
           ggtheme = theme_bw(),risk.table = 0.25, 
           palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media',break.x.by = 2,
           title = 'OP50') 
```



Shows errorbars and lineplot based on one of Anders papers https://onlinelibrary.wiley.com/doi/full/10.1111/acel.12165
This is done by extracting the surviving proportion, standard errors and time from survfit().  and adding a startpoint where the survival is 100% at 0 hours.


Now we test for difference between the curves with both log-rank and gehan-wilcoxon and with an cox proportional hazard model

```{r}
data$Survobj <- with(data,Surv(data$Time, event = data$status))
survdiff(Survobj ~media + mutant, data = data, rho = 0) 
survdiff(Survobj ~media + mutant, data = data, rho = 1) 
fit_both<- coxph(Survobj ~media + mutant, data = data)
summary(fit_both)
fit_int <- coxph(Survobj ~media + mutant + mutant:media, data = data)
AIC(fit_both)
fit_med <- coxph(Survobj ~media, data = data)
summary(fit_med)
AIC(fit_med)
fit_mut <- coxph(Survobj ~mutant, data = data)
summary(fit_mut)
AIC(fit_mut)
```
Then we compare the different groups individually

```{r}
d_OP50 <- data[data$media == 'OP50',]
d_13 <- data[data$media == "FG13",]
d_20 <- data[data$media == "FG20",]
d_fly <- rbind(d_13, d_20)
d_2050 <- rbind(d_20, d_OP50)
d_1350 <- rbind(d_13, d_OP50)
```

First tests for each combination first fly media

```{r}
survdiff(Surv(Time,status) ~ media, data = d_fly, rho = 0)
survdiff(Surv(Time,status) ~ media, data = d_fly, rho = 1)
fit <- coxph(Surv(Time,status) ~media, data = d_fly)
summary(fit)
```

Then 20 degrees and OP50
```{r}
survdiff(Surv(Time,status) ~ media, data = d_2050, rho = 0)
survdiff(Surv(Time,status) ~ media, data = d_2050, rho = 1)
fit <- coxph(Surv(Time,status) ~media, data = d_2050)
summary(fit)
```

Then for 13 degrees and OP50
```{r}
survdiff(Surv(Time,status) ~ media, data = d_1350, rho = 0)
survdiff(Surv(Time,status) ~ media, data = d_1350, rho = 1)
fit <- coxph(Surv(Time,status) ~media, data = d_1350)
summary(fit)
```

 Where only 20 degrees and 13 degrees are significantly different from each other. But there is also some difference between 20 degrees and OP50. Where there isn´t a big difference between 13 degrees and
 OP50.
 
 Then dose response curves with first the two fly media.
 
```{r}
plot(df_fly_13$Time[2:9],df_fly_13$Surv[2:9], pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival heat stress for'~italic(C.elegans)),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
lines(df_fly_13$Time,df_fly_13$Surv)
arrows(df_fly_13$Time, df_fly_13$Surv-df_fly_13$Std.error, df_fly_13$Time, 
       df_fly_13$Surv+df_fly_13$Std.error, length=0.05, angle=90, code=3, col = 'black')
points(df_fly_20$Time[2:9],df_fly_20$Surv[2:9], pch = 6)
lines(df_fly_20$Time,df_fly_20$Surv,lty = 3)
arrows(df_fly_20$Time, df_fly_20$Surv-df_fly_20$Std.error, df_fly_20$Time, 
       df_fly_20$Surv+df_fly_20$Std.error, length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,6),lty = c(1,3), 
       legend = c(expression('Fly gut 13'*degree*C),
                  expression('Fly gut 20'*degree*C)), bty = 'n')
```

Than 20 degrees and OP50

```{r}
plot(df_fly_20$Time[2:9],df_fly_20$Surv[2:9], pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival heat stress for'~italic(C.elegans)),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
points(df_OP50$Time[2:9],df_OP50$Surv[2:9], pch = 1, cex = 1.2)
lines(df_OP50$Time,df_OP50$Surv, lty = 2)
arrows(df_OP50$Time, df_OP50$Surv-df_OP50$Std.error,
      df_OP50$Time, df_OP50$Surv+df_OP50$Std.error, 
      length=0.05, angle=90, code=3, col = 'black')
lines(df_fly_20$Time,df_fly_20$Surv,lty = 1)
arrows(df_fly_20$Time, df_fly_20$Surv-df_fly_20$Std.error, df_fly_20$Time, 
       df_fly_20$Surv+df_fly_20$Std.error, length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,1),lty = c(1,2), 
       legend = c(expression('Fly gut 20'*degree*C),
                  expression(italic(E.coli) ~ OP50)), bty = 'n')
```

Then 13 degrees and OP50
```{r}
plot(df_fly_13$Time[2:9],df_fly_13$Surv[2:9], pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival heat stress for'~italic(C.elegans)),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
lines(df_fly_13$Time,df_fly_13$Surv)
arrows(df_fly_13$Time, df_fly_13$Surv-df_fly_13$Std.error, df_fly_13$Time, 
       df_fly_13$Surv+df_fly_13$Std.error, length=0.05, angle=90, code=3, col = 'black')
points(df_OP50$Time[2:9],df_OP50$Surv[2:9], pch = 1, cex = 1.2)
lines(df_OP50$Time,df_OP50$Surv, lty = 2)
arrows(df_OP50$Time, df_OP50$Surv-df_OP50$Std.error,
      df_OP50$Time, df_OP50$Surv+df_OP50$Std.error, 
      length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,1),lty = c(1,2), 
       legend = c(expression('Fly gut 13'*degree*C),
                  expression(italic(E.coli) ~ OP50)), bty = 'n')
```
Then Kaplan meier curves
```{r}
km_fly <- survfit(Surv(Time,status) ~ media, data = d_fly,conf.type = "log-log", error = "greenwood")
km_2050 <- survfit(Surv(Time,status) ~ media, data = d_2050,conf.type = "log-log", error = "greenwood")
km_1350 <- survfit(Surv(Time,status) ~ media, data = d_1350,conf.type = "log-log", error = "greenwood")
ggsurvplot(km_fly,data = d_fly, conf.int = TRUE,
           ggtheme = theme_bw(),risk.table = 0.25,
           palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media',break.x.by = 2)
ggsurvplot(km_2050,data = d_2050, conf.int = TRUE, 
          ggtheme = theme_bw(), risk.table = 0.25,
          palette ='aaas', surv.median.line = 'hv',
          legend.title = 'Media',break.x.by = 2)
ggsurvplot(km_1350,data = d_1350, conf.int = TRUE, 
           ggtheme = theme_bw(),risk.table = 0.25,
           palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media', break.x.by = 2) 
```

