---
title: "Analysis_ramping"
author: "Jonas Gehrlein"
date: "29 okt 2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(survival) 
library(here)
library(survminer)
```

First we download the data and remove embty columns
```{r}
data <- read.table(here('data','Ramp_4days.csv'), header = TRUE, 
                   sep = ";", dec = ",")
head(data)
data <- data[,1:3]
data <- data[complete.cases(data),]
```

here() should show that your position is in the folder 7.semester else either download
7.semester from github again or create a .Rproj file in the folder on your computer
One fly died from water entering the tube and is marked NA so it is removed.
Or make a new .Rproj file in the folder 7.semester. There is a problem with the numbers of the  so we change that with col.names()
```{r}
colnames(data)<-c('ID',"Ctmax","Temp")
head(data)
```
We then compare the two acclimation temperatures first with a boxplot
```{r}
boxplot(data$Ctmax ~ data$Temp, xlab = 'Acclimation Temperature',
        ylab = 'CTmax', 
        main = expression('CTmax for' ~italic(D.melanogaster) ~ 'acclimated to 13 or 20 degrees' ))
```
 This shows a difference in median Ctmax. To look more finegrained at
 the distribution of CTmax we use a survival curve
 

Then we create a survival object with the survival package and make a kaplan-meier curve
```{r}
data$status <- rep(1,75)
data$Survobj <- with(data,Surv(data$Ctmax, event = data$status))
km <- survfit(Survobj ~ Temp, data = data,conf.type = "log-log")
```
Which shows a clear difference

```{r}
survdiff(Survobj ~Temp, data = data, rho = 0) 
survdiff(Survobj ~Temp, data = data, rho = 1) 
fit <- coxph(Surv(Ctmax,status) ~Temp, data = data)
summary(fit)
```

Which both find a significant difference between the groups 
But we can make a nicer survival curve with a different package survminer 
and add readable confidence intervals

```{r}
mycols <- c(gray.colors(4,0,0.9))
ggsurvplot(km,data = data, conf.int = TRUE, ggtheme = theme_bw(),
           xlim = c(37,42), xlab = 'Temperature (°C)',ylab = 'Surviving fraction',title = expression('                                Survival of' ~ italic(D.) ~ italic(melanogaster)),palette = mycols,
           legend = c(0.8,0.86), legend.title = 'Developmental temperature',
           legend.labs = c('13°C','20°C'),
           break.x.by = 1, surv.median.line = 'hv')
```

Which very clearly shows the difference between the different acclimation temperatures
