---
title: "Analysis_ramping"
author: "Jonas Gehrlein"
date: "29 okt 2018"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(survival) 
library(here)
library(survminer)
```
code for fitting Kaplan-Meier and log-rank test and for displaying 
survival curves for each type of media

```{r}
data <- read.table(here('data','Ramp_4days.csv'), header = TRUE, sep = ";", dec = ",")
head(data)
data <- data[,1:3]
data <- data[complete.cases(data),]
```

here() should show that your position is in the folder 7.semester else either download
7.semester from github again or create a .Rproj file in the folder on your computer
One fly died from water entering the tube and is marked NA so it is removed.
Or make a new .Rproj file in the folder 7.semester. There is a problem with the numbers of the  so we change that with col.names()
```{r}
colnames(data)<-c('ID',"Ctmax","Temp")
head(data)
```
We then compare the two acclimation temperatures first with a boxplot
```{r}
boxplot(data$Ctmax ~ data$Temp)
```
 This shows basically no difference between the groups. 
 We try to see what is happening with a survival curve
 

Then we create a survival object with the survival package and make a kaplan-meier curve
```{r}
data$status <- rep(1,75)
data$Survobj <- with(data,Surv(data$Ctmax, event = data$status))
km <- survfit(Survobj ~ Temp, data = data,conf.type = "log-log")
plot(km, col = c("blue","red"), xlim = c(30,42) ) 
legend("bottomleft",legend = c("13","20" ), col = c("blue","red"), lwd=1)
```
Which also don`t show any meaningful difference between the groups
Now we test for difference between the curves with both log-rank and gehan-wilcoxon

```{r}
survdiff(Survobj ~Temp, data = data, rho = 0) 
survdiff(Survobj ~Temp, data = data, rho = 1) 
```

Which both don`t find a significant difference between the groups 
But we can make a nicer survival curve with a different package survminer 
and add readable confidence intervals

```{r}
ggsurvplot(km,data = data, conf.int = TRUE, ggtheme = theme_bw(),
           xlim = c(30,42), xlab = "CTmax")
```

Whic very clearly shows that the problem is that the two lines are far to close 
together.
