---
title: "Survival of C.elegans with three different media"
author: "Jonas Gehrlein"
date: "12 nov 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(survival) 
library(here)
library(survminer)
```
code for fitting Kaplan-Meier and log-rank test and for displaying 
survival curves for each type of media and mutant.

```{r}
data <- read.table(here('data','Worm_mutant.csv'), header = TRUE, sep = ";", dec = ",")
head(data)
```

here() should show that your position is in the folder 7.semester else select the active project to be 7.semester in the upper right corner. 

Or make a new .Rproj file in the folder 7.semester. There is a problem with the numbers of the worms so we change that with col.names()
```{r}
colnames(data)<-c("Time","status",'Replicate', 'media',
                  'mutant')
head(data)
```

Then we create a survival object with the survival package and make a kaplan-meier curve

First we see for each mutant
```{r}
d_RRF <- subset(data, mutant == 'RRF')
d_PMK <- subset(data,mutant == 'PMK')
d_DAF <- subset(data, mutant == 'DAF')
```
```{r}
d_RRF$Survobj <- with(d_RRF,Surv(d_RRF$Time, event = d_RRF$status))
km_RRF <- survfit(Survobj ~ media, data = d_RRF,conf.type = "log-log", error = "greenwood")
s_km_RRF <- summary(km_RRF)
ggsurvplot(km_RRF,data = d_RRF, conf.int = TRUE, 
           ggtheme = theme_bw(),risk.table = 0.25,
           palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media',break.x.by = 2,
           title = 'RRF-3') 
```

```{r}
df_fly_13_r <- data.frame(c(0,s_km_RRF$time[1:6]),c(1,s_km_RRF$surv[1:6]),                        c(0,s_km_RRF$std.err[1:6]))
colnames(df_fly_13_r) <- c('Time','Surv','Std.error')
df_fly_20_r <- data.frame(c(0,s_km_RRF$time[7:11])
                        ,c(1,s_km_RRF$surv[7:11]),    c(0,s_km_RRF$std.err[7:11]))
colnames(df_fly_20_r) <- c('Time','Surv','Std.error')
df_OP50_r <- data.frame(c(0,s_km_RRF$time[12:17])
                      ,c(1,s_km_RRF$surv[12:17]),
                        c(0,s_km_RRF$std.err[12:17]))
colnames(df_OP50_r) <- c('Time','Surv','Std.error')

plot(df_fly_13_r$Time[2:7],df_fly_13_r$Surv[2:7], pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival heat stress for'~italic(C.elegans)),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
lines(df_fly_13_r$Time,df_fly_13_r$Surv)
arrows(df_fly_13_r$Time, df_fly_13_r$Surv-df_fly_13_r$Std.error, df_fly_13_r$Time, 
       df_fly_13_r$Surv+df_fly_13_r$Std.error, length=0.05, angle=90, code=3, col = 'black')
points(df_OP50_r$Time[2:9],df_OP50_r$Surv[2:9], pch = 1, cex = 1.2)
lines(df_OP50_r$Time,df_OP50_r$Surv, lty = 2)
arrows(df_OP50_r$Time, df_OP50_r$Surv-df_OP50_r$Std.error,
      df_OP50_r$Time, df_OP50_r$Surv+df_OP50_r$Std.error, 
      length=0.05, angle=90, code=3, col = 'black')
points(df_fly_20_r$Time[2:9],df_fly_20_r$Surv[2:9], pch = 6)
lines(df_fly_20_r$Time,df_fly_20_r$Surv,lty = 3)
arrows(df_fly_20_r$Time, df_fly_20_r$Surv-df_fly_20_r$Std.error, df_fly_20_r$Time,df_fly_20_r$Surv+df_fly_20_r$Std.error, length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,6,1),lty = c(1,3,2), 
       legend = c(expression('Fly gut 13'*degree*C),
                  expression('Fly gut 20'*degree*C),
                  expression(italic(E.coli) ~ OP50)), bty = 'n')
```

```{r}
d_PMK$Survobj <- with(d_PMK,Surv(d_PMK$Time, event = d_PMK$status))
km_PMK <- survfit(Survobj ~ media, data = d_PMK,conf.type = "log-log", error = "greenwood")
s_km_PMK <- summary(km_PMK)
ggsurvplot(km_PMK,data = d_PMK, conf.int = TRUE, 
           ggtheme = theme_bw(),risk.table = 0.25, 
           palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media',break.x.by = 2,
           title = 'PMK-1') 
```

```{r}
df_fly_13_P <- data.frame(c(0,s_km_PMK$time[1:5]),
                        c(1,s_km_PMK$surv[1:5]),
                        c(0,s_km_PMK$std.err[1:5]))
colnames(df_fly_13_P) <- c('Time','Surv','Std.error')
df_fly_20_P <- data.frame(c(0,s_km_PMK$time[6:10]),
                        c(1,s_km_PMK$surv[6:10]),
                        c(0,s_km_PMK$std.err[6:10]))
colnames(df_fly_20_P) <- c('Time','Surv','Std.error')
df_OP50_P <- data.frame(c(0,s_km_PMK$time[11:17]),
                      c(1,s_km_PMK$surv[11:17]),
                        c(0,s_km_PMK$std.err[11:17]))
colnames(df_OP50_P) <- c('Time','Surv','Std.error')

plot(df_fly_13_P$Time[2:9],df_fly_13_P$Surv[2:9], pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival heat stress for'~italic(C.elegans)),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
lines(df_fly_13_P$Time,df_fly_13_P$Surv)
arrows(df_fly_13_P$Time, df_fly_13_P$Surv-df_fly_13_P$Std.error, df_fly_13_P$Time,df_fly_13_P$Surv+df_fly_13_P$Std.error, length=0.05, angle=90, code=3, col = 'black')
points(df_OP50_P$Time[2:9],df_OP50_P$Surv[2:9], pch = 1, cex = 1.2)
lines(df_OP50_P$Time,df_OP50_P$Surv, lty = 2)
arrows(df_OP50_P$Time, df_OP50_P$Surv-df_OP50_P$Std.error,
      df_OP50_P$Time, df_OP50_P$Surv+df_OP50_P$Std.error, 
      length=0.05, angle=90, code=3, col = 'black')
points(df_fly_20_P$Time[2:9],df_fly_20_P$Surv[2:9], pch = 6)
lines(df_fly_20_P$Time,df_fly_20_P$Surv,lty = 3)
arrows(df_fly_20_P$Time, df_fly_20_P$Surv-df_fly_20_P$Std.error, df_fly_20_P$Time,df_fly_20_P$Surv+df_fly_20_P$Std.error, length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,6,1),lty = c(1,3,2), 
       legend = c(expression('Fly gut 13'*degree*C),
                  expression('Fly gut 20'*degree*C),
                  expression(italic(E.coli) ~ OP50)), bty = 'n')
```

```{r}
d_DAF$Survobj <- with(d_DAF,Surv(d_DAF$Time, event = d_DAF$status))
km_DAF <- survfit(Survobj ~ media, data = d_DAF,conf.type = "log-log", error = "greenwood")
s_km_DAF <- summary(km_DAF)
ggsurvplot(km_DAF,data = d_DAF, conf.int = TRUE, 
           ggtheme = theme_bw(),risk.table = 0.25,
           palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media',break.x.by = 2,
           title = 'DAF-16') 
```

```{r}
df_fly_13_d <- data.frame(c(0,s_km_DAF$time[1:3]),
                        c(1,s_km_DAF$surv[1:3]),
                        c(0,s_km_DAF$std.err[1:3]))
colnames(df_fly_13_d) <- c('Time','Surv','Std.error')
df_fly_20_d <- data.frame(c(0,s_km_DAF$time[4:8]),
                        c(1,s_km_DAF$surv[4:8]),
                        c(0,s_km_DAF$std.err[4:8]))
colnames(df_fly_20_d) <- c('Time','Surv','Std.error')
df_OP50_d <- data.frame(c(0,s_km_DAF$time[9:15]),
                      c(1,s_km_DAF$surv[9:15]),
                        c(0,s_km_DAF$std.err[9:15]))
colnames(df_OP50_d) <- c('Time','Surv','Std.error')

plot(df_fly_13_d$Time[2:9],df_fly_13_d$Surv[2:9], pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival heat stress for'~italic(C.elegans)),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
lines(df_fly_13_d$Time,df_fly_13_d$Surv)
arrows(df_fly_13_d$Time, df_fly_13_d$Surv-df_fly_13_d$Std.error, df_fly_13_d$Time,df_fly_13_d$Surv+df_fly_13_d$Std.error, length=0.05, angle=90, code=3, col = 'black')
points(df_OP50_d$Time[2:9],df_OP50_d$Surv[2:9], pch = 1, cex = 1.2)
lines(df_OP50_d$Time,df_OP50_d$Surv, lty = 2)
arrows(df_OP50_d$Time, df_OP50_d$Surv-df_OP50_d$Std.error,
      df_OP50_d$Time, df_OP50_d$Surv+df_OP50_d$Std.error, 
      length=0.05, angle=90, code=3, col = 'black')
points(df_fly_20_d$Time[2:9],df_fly_20_d$Surv[2:9], pch = 6)
lines(df_fly_20_d$Time,df_fly_20_d$Surv,lty = 3)
arrows(df_fly_20_d$Time, df_fly_20_d$Surv-df_fly_20_d$Std.error, df_fly_20_d$Time,df_fly_20_d$Surv+df_fly_20_d$Std.error, length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,6,1),lty = c(1,3,2), 
       legend = c(expression('Fly gut 13'*degree*C),
                  expression('Fly gut 20'*degree*C),
                  expression(italic(E.coli) ~ OP50)), bty = 'n')
```
Then we compare by medium
```{r}
data_OP50 <- subset(data, media == 'OP50')
data_FG13 <- subset(data, media == 'FG13')
data_FG20 <- subset(data, media == 'FG20')
```

```{r}
data_OP50$Survobj <- with(data_OP50,Surv(data_OP50$Time, event = data_OP50$status))
km_OP50 <- survfit(Survobj ~ mutant, data = data_OP50,conf.type = "log-log", error = "greenwood")
s_km_OP50 <- summary(km_OP50)
ggsurvplot(km_OP50,data = data_OP50, conf.int = TRUE, 
           ggtheme = theme_bw(),risk.table = 0.25, 
           palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media',break.x.by = 2,
           title = 'OP50') 
```


```{r}
data_FG13$Survobj <- with(data_FG13,Surv(data_FG13$Time, event = data_FG13$status))
km_FG13 <- survfit(Survobj ~ mutant, data = data_FG13,conf.type = "log-log", error = "greenwood")
s_km_FG13 <- summary(km_FG13)
ggsurvplot(km_FG13,data = data_FG13, conf.int = TRUE, 
           ggtheme = theme_bw(),risk.table = 0.25, 
           palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media',break.x.by = 2,
           title = 'FG13') 
```

```{r}
data_FG20$Survobj <- with(data_FG20,Surv(data_FG20$Time, event = data_FG20$status))
km_FG20 <- survfit(Survobj ~ mutant, data = data_FG20,conf.type = "log-log", error = "greenwood")
s_km_FG20 <- summary(km_FG20)
ggsurvplot(km_FG20,data = data_FG20, conf.int = TRUE, 
           ggtheme = theme_bw(),risk.table = 0.25, 
           palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media',break.x.by = 2,
           title = 'OP50') 
```



Shows errorbars and lineplot based on one of Anders papers https://onlinelibrary.wiley.com/doi/full/10.1111/acel.12165
This is done by extracting the surviving proportion, standard errors and time from survfit().  and adding a startpoint where the survival is 100% at 0 hours.


Now we test for difference between the curves with both log-rank and gehan-wilcoxon and with an cox proportional hazard model

```{r}
data$Survobj <- with(data,Surv(data$Time, event = data$status))
survdiff(Survobj ~media + mutant, data = data, rho = 0) 
survdiff(Survobj ~media + mutant, data = data, rho = 1) 
fit_both<- coxph(Survobj ~media + mutant, data = data)
summary(fit_both)
AIC(fit_both)
fit_int <- coxph(Survobj ~media + mutant + mutant:media, data = data)
AIC(fit_int)
fit_med <- coxph(Survobj ~media, data = data)
summary(fit_med)
AIC(fit_med)
fit_mut <- coxph(Survobj ~mutant, data = data)
summary(fit_mut)
AIC(fit_mut)
```

Best model must include both media and mutant strain but not necessarily an interaction between them. This could be driven by few observations for FG20 in RRF-3
Then we do a check on the proportional hazard assumption
```{r}
prop_both <- cox.zph(fit_both)
prop_int <- cox.zph(fit_int)
plot(prop_both)
plot(prop_int)
```
Both schoenfield residuals and the test is not significant so we proceed as if the assumptions is true for our dataset.

Then we plot predictions for both models
```{r}
plot(survfit(fit_both, newdata = data.frame(mutant = 'RRF', media = 'OP50')), lty = 2, xaxp = c(0,18,18), col = 'blue',
     main = 'Predictions for model without interactions OP50')
lines(survfit(fit_both, newdata = data.frame(mutant = 'PMK', media = 'OP50')), col = 'firebrick')
lines(survfit(fit_both, newdata = data.frame(mutant = 'DAF', media = 'OP50')), lty =3)
legend('topright', legend = c('rrf-3','pmk-1','daf-16'), 
        lty = c(2,1,3), col = c('blue','firebrick','black'), 
        bty = 'n')
plot(survfit(fit_int, newdata = data.frame(mutant = 'RRF', media = 'OP50')), lty = 2, xaxp = c(0,18,18), col = 'blue',
     main = 'Predictions for model with interactions OP50')
lines(survfit(fit_int, newdata = data.frame(mutant = 'PMK', media = 'OP50')), col = 'firebrick')
lines(survfit(fit_int, newdata = data.frame(mutant = 'DAF', media = 'OP50')), lty =3)
legend('topright', legend = c('rrf-3','pmk-1','daf-16'), 
        lty = c(2,1,3), col = c('blue','firebrick','black'), 
        bty = 'n')
plot(survfit(fit_both, newdata = data.frame(mutant = 'RRF', media = 'FG13')), lty = 2, xaxp = c(0,18,9), col = 'blue',
     main = 'Predictions for model without interactions FG13',
     xlab = 'hours heat stress')
lines(survfit(fit_both, newdata = data.frame(mutant = 'PMK', media = 'FG13')), col = 'firebrick')
lines(survfit(fit_both, newdata = data.frame(mutant = 'DAF', media = 'FG13')), lty =3)
legend('topright', legend = c('rrf-3','pmk-1','daf-16'), 
        lty = c(2,1,3), col = c('blue','firebrick','black'), 
        bty = 'n')
plot(survfit(fit_int, newdata = data.frame(mutant = 'RRF', media = 'FG13')), lty = 2, xaxp = c(0,18,18), col = 'blue',
     main = 'Predictions for model with interactions FG13')
lines(survfit(fit_int, newdata = data.frame(mutant = 'PMK', media = 'FG13')), col = 'firebrick')
lines(survfit(fit_int, newdata = data.frame(mutant = 'DAF', media = 'FG13')), lty =3)
legend('topright', legend = c('rrf-3','pmk-1','daf-16'), 
        lty = c(2,1,3), col = c('blue','firebrick','black'), 
        bty = 'n')
plot(survfit(fit_both, newdata = data.frame(mutant = 'RRF', media = 'FG20')), lty = 2, xaxp = c(0,18,9), col = 'blue',
     main = 'Predictions for model without interactions FG20',
     xlab = 'hours heat stress')
lines(survfit(fit_both, newdata = data.frame(mutant = 'PMK', media = 'FG20')), col = 'firebrick')
lines(survfit(fit_both, newdata = data.frame(mutant = 'DAF', media = 'FG20')), lty =3)
legend('topright', legend = c('rrf-3','pmk-1','daf-16'), 
        lty = c(2,1,3), col = c('blue','firebrick','black'), 
        bty = 'n')
plot(survfit(fit_int, newdata = data.frame(mutant = 'RRF', media = 'FG20')), lty = 2, xaxp = c(0,18,18), col = 'blue',
     main = 'Predictions for model with interactions FG20')
lines(survfit(fit_int, newdata = data.frame(mutant = 'PMK', media = 'FG20')), col = 'firebrick')
lines(survfit(fit_int, newdata = data.frame(mutant = 'DAF', media = 'FG20')), lty =3)
legend('topright', legend = c('rrf-3','pmk-1','daf-16'), 
        lty = c(2,1,3), col = c('blue','firebrick','black'), 
        bty = 'n')

```
We check how well predictions line up with reality

```{r}
plot(survfit(fit_both, newdata = data.frame(mutant = 'RRF', media = 'OP50')), lty = 2, xaxp = c(0,18,9), col = 'blue',
     main = 'Predictions for model without interactions OP50')
points(df_OP50_r$Time,df_OP50_r$Surv, col = 'blue', pch = 16)
lines(survfit(fit_both, newdata = data.frame(mutant = 'DAF', media = 'OP50')), lty = 2, col = 'black',)
points(df_OP50_d$Time,df_OP50_d$Surv, col = 'black', pch = 16)
lines(survfit(fit_both, newdata = data.frame(mutant = 'PMK', media = 'OP50')), lty = 2, col = 'firebrick',)
points(df_OP50_P$Time,df_OP50_P$Surv, col = 'firebrick', pch = 16)
legend('topright', legend = c('rrf-3','pmk-1','daf-16'), 
        lty = c(2,1,3), col = c('blue','firebrick','black'), 
        bty = 'n', pch = 16)
```
Fit is very bad as is clear both concordance and R^2 from the model.
But there is far too little uncertainty in the confidence intervals. Which 
overestimates survival heavily for this group. Lets try the other model

```{r}
plot(survfit(fit_int, newdata = data.frame(mutant = 'RRF', media = 'OP50')), lty = 2, xaxp = c(0,18,9), col = 'blue',
     main = 'Predictions for model with interactions OP50')
points(df_OP50_r$Time,df_OP50_r$Surv, col = 'blue', pch = 16)
lines(survfit(fit_int, newdata = data.frame(mutant = 'DAF', media = 'OP50')), lty = 2, col = 'black',)
points(df_OP50_d$Time,df_OP50_d$Surv, col = 'black', pch = 16)
lines(survfit(fit_int, newdata = data.frame(mutant = 'PMK', media = 'OP50')), lty = 2, col = 'firebrick',)
points(df_OP50_P$Time,df_OP50_P$Surv, col = 'firebrick', pch = 16)
legend('topright', legend = c('rrf-3','pmk-1','daf-16'), 
        lty = c(2,1,3), col = c('blue','firebrick','black'), 
        bty = 'n', pch = 16)
```

 Still on the low end but a lot better. Seems that without interactions it overestimates survival on OP50.
```{r}
plot(survfit(fit_both, newdata = data.frame(mutant = 'RRF', media = 'FG13')), lty = 2, xaxp = c(0,18,9), col = 'blue',
     main = 'Predictions for model with interactions FG13')
points(df_fly_13_r$Time,df_fly_13_r$Surv, col = 'blue', pch = 16)
lines(survfit(fit_both, newdata = data.frame(mutant = 'DAF', media = 'FG13')), lty = 2, col = 'black',)
points(df_fly_13_d$Time,df_fly_13_d$Surv, col = 'black', pch = 16)
lines(survfit(fit_both, newdata = data.frame(mutant = 'PMK', media = 'FG13')), lty = 2, col = 'firebrick',)
points(df_fly_13_P$Time,df_fly_13_P$Surv, col = 'firebrick', pch = 16)
legend('topright', legend = c('rrf-3','pmk-1','daf-16'), 
        lty = c(2,1,3), col = c('blue','firebrick','black'), 
        bty = 'n', pch = 16)
```

Fits notably better 

```{r}
plot(survfit(fit_int, newdata = data.frame(mutant = 'RRF', media = 'FG13')), lty = 2, xaxp = c(0,18,9), col = 'blue',
     main = 'Predictions for model with interactions FG13')
points(df_fly_13_r$Time,df_fly_13_r$Surv, col = 'blue', pch = 16)
lines(survfit(fit_int, newdata = data.frame(mutant = 'DAF', media = 'FG13')), lty = 2, col = 'black',)
points(df_fly_13_d$Time,df_fly_13_d$Surv, col = 'black', pch = 16)
lines(survfit(fit_int, newdata = data.frame(mutant = 'PMK', media = 'FG13')), lty = 2, col = 'firebrick',)
points(df_fly_13_P$Time,df_fly_13_P$Surv, col = 'firebrick', pch = 16)
legend('topright', legend = c('rrf-3','pmk-1','daf-16'), 
        lty = c(2,1,3), col = c('blue','firebrick','black'), 
        bty = 'n', pch = 16)
```

Change opacity on some points to fix overlap
```{r}
plot(survfit(fit_both, newdata = data.frame(mutant = 'RRF', media = 'FG20')), lty = 2, xaxp = c(0,18,9), col = 'blue',
     main = 'Predictions for model without interactions FG20')
points(df_fly_20_r$Time,df_fly_20_r$Surv, col = 'blue', pch = 16)
lines(survfit(fit_both, newdata = data.frame(mutant = 'DAF', media = 'FG20')), lty = 2, col = 'black',)
points(df_fly_20_d$Time,df_fly_20_d$Surv, col = 'black', pch = 16)
lines(survfit(fit_both, newdata = data.frame(mutant = 'PMK', media = 'FG20')), lty = 2, col = 'firebrick',)
points(df_fly_20_P$Time,df_fly_20_P$Surv, col = 'firebrick', pch = 16)
legend('topright', legend = c('rrf-3','pmk-1','daf-16'), 
        lty = c(2,1,3), col = c('blue','firebrick','black'), 
        bty = 'n', pch = 16)
```

Fitsvery bad with far too much uncertainty

```{r}
plot(survfit(fit_int, newdata = data.frame(mutant = 'RRF', media = 'FG20')), lty = 2, xaxp = c(0,18,9), col = 'blue',
     main = 'Predictions for model with interactions FG20')
points(df_fly_20_r$Time,df_fly_20_r$Surv, col = 'blue', pch = 16)
lines(survfit(fit_int, newdata = data.frame(mutant = 'DAF', media = 'FG20')), lty = 2, col = 'black',)
points(df_fly_20_d$Time,df_fly_20_d$Surv, col = 'black', pch = 16)
lines(survfit(fit_int, newdata = data.frame(mutant = 'PMK', media = 'FG13')), lty = 2, col = 'firebrick',)
points(df_fly_20_P$Time,df_fly_20_P$Surv, col = 'firebrick', pch = 16)
legend('topright', legend = c('rrf-3','pmk-1','daf-16'), 
        lty = c(2,1,3), col = c('blue','firebrick','black'), 
        bty = 'n', pch = 16)
```

The model has a problem with catching that all worms die around the same time and may therefore overestimate survival as it fits okay until the worms start 
dying. Some other model structure is maybe necessary or not all important covariates of survival are measured.


