---
title: "Survival"
author: "Jonas Gehrlein"
date: "24 okt 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(survival) 
library(here)
library(survminer)
```
code for fitting Kaplan-Meier and log-rank test and for displaying 
survival curves for each type of media

```{r}
here()
data <- read.table(here('data','Trial_worm.csv'), header = TRUE, sep = ";", dec = ",")
head(data)
```

here() should show that your position is in the folder 7.semester else select the active project to be 7.semester in the upper right corner. 

Or make a new .Rproj file in the folder 7.semester. There is a problem with the numbers of the worms so we change that with col.names()
```{r}
colnames(data)<-c('worm_nr',"Time","status","Replicate","media")
head(data)
```

Then we create a survival object with the survival package and make a kaplan-meier curve
```{r}
data$Survobj <- with(data,Surv(data$Time, event = data$status))
km <- survfit(Survobj ~ media, data = data,conf.type = "log-log")
plot(km, col = c("blue","red") ) 
legend("bottomleft",legend = c("FG20","OP50" ), col = c("blue","red"), lwd=1)
```

Now we test for difference between the curves with both log-rank and gehan-wilcoxon

```{r}
survdiff(Survobj ~media, data = data, rho = 0) 
survdiff(Survobj ~media, data = data, rho = 1) 
```

Which both don`t find a significant difference between the groups But we can make a nicer survival curve with a different package survminer and add readable confidence intervals

```{r}
ggsurvplot(km,data = data, conf.int = TRUE, ggtheme = theme_bw(),
           risk.table = 0.25)
```

Which shows that the problem is the uncertainty in OP50 because of not enough samples

