---
title: "Survival of C.elegans with three different media"
author: "Jonas Gehrlein"
date: "12 nov 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(survival) 
library(here)
library(survminer)
```
code for fitting Kaplan-Meier and log-rank test and for displaying 
survival curves for each type of media

```{r}
data <- read.table(here('data','Worm_4days.csv'), header = TRUE, sep = ";", dec = ",")
head(data)
```

here() should show that your position is in the folder 7.semester else select the active project to be 7.semester in the upper right corner. 

Or make a new .Rproj file in the folder 7.semester. There is a problem with the numbers of the worms so we change that with col.names()
```{r}
colnames(data)<-c("Time","status","media",'Replicate')
head(data)
```

Then we create a survival object with the survival package and make a kaplan-meier curve
```{r}
data$Survobj <- with(data,Surv(data$Time, event = data$status))
km <- survfit(Survobj ~ media, data = data,conf.type = "log-log", error = "greenwood")
s_km <- summary(km)
plot(km, col = c("red","black",'blue'))
fit <- coxph(Survobj ~ media, data = data)
summary(fit)
```

Shows errorbars and lineplot based on one of Anders papers https://onlinelibrary.wiley.com/doi/full/10.1111/acel.12165
This is done by extracting the surviving proportion, standard errors and time from survfit().  and adding a startpoint where the survival is 100% at 0 hours.

```{r}
df_fly_13 <- data.frame(c(0,s_km$time[1:8]),c(1,s_km$surv[1:8]),
                        c(0,s_km$std.err[1:8]))
colnames(df_fly_13) <- c('Time','Surv','Std.error')
df_fly_20 <- data.frame(c(0,s_km$time[9:16]),c(1,s_km$surv[9:16]),
                        c(0,s_km$std.err[9:16]))
colnames(df_fly_20) <- c('Time','Surv','Std.error')
df_OP50 <- data.frame(c(0,s_km$time[17:24]),c(1,s_km$surv[17:24]),
                        c(0,s_km$std.err[17:24]))
colnames(df_OP50) <- c('Time','Surv','Std.error')

plot(df_fly_13$Time[2:9],df_fly_13$Surv[2:9], pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival heat stress for'~italic(C.elegans)),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
lines(df_fly_13$Time,df_fly_13$Surv)
arrows(df_fly_13$Time, df_fly_13$Surv-df_fly_13$Std.error, df_fly_13$Time, 
       df_fly_13$Surv+df_fly_13$Std.error, length=0.05, angle=90, code=3, col = 'black')
points(df_OP50$Time[2:9],df_OP50$Surv[2:9], pch = 1, cex = 1.2)
lines(df_OP50$Time,df_OP50$Surv, lty = 2)
arrows(df_OP50$Time, df_OP50$Surv-df_OP50$Std.error,
      df_OP50$Time, df_OP50$Surv+df_OP50$Std.error, 
      length=0.05, angle=90, code=3, col = 'black')
points(df_fly_20$Time[2:9],df_fly_20$Surv[2:9], pch = 6)
lines(df_fly_20$Time,df_fly_20$Surv,lty = 3)
arrows(df_fly_20$Time, df_fly_20$Surv-df_fly_20$Std.error, df_fly_20$Time, 
       df_fly_20$Surv+df_fly_20$Std.error, length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,6,1),lty = c(1,3,2), 
       legend = c(expression('Fly gut 13'*degree*C),
                  expression('Fly gut 20'*degree*C),
                  expression(italic(E.coli) ~ OP50)), bty = 'n')
```

Then we create a normal Kaplan-Meier curve
```{r}
ggsurvplot(km,data = data, conf.int = TRUE, ggtheme = theme_bw(),
           risk.table = 0.25, palette ='aaas', surv.median.line = 'hv',legend.title = 'Media',break.x.by = 2) 
```

Now we test for difference between the curves with both log-rank and gehan-wilcoxon and with an cox proportional hazard model

```{r}
survdiff(Survobj ~media, data = data, rho = 0) 
survdiff(Survobj ~media, data = data, rho = 1) 
fit <- coxph(Survobj ~media, data = data)
summary(fit)
```

Then we compare the different groups individually

```{r}
d_OP50 <- data[data$media == 'OP50',]
d_13 <- data[data$media == "FG13",]
d_20 <- data[data$media == "FG20",]
d_fly <- rbind(d_13, d_20)
d_2050 <- rbind(d_20, d_OP50)
d_1350 <- rbind(d_13, d_OP50)
```

First tests for each combination first fly media

```{r}
survdiff(Surv(Time,status) ~ media, data = d_fly, rho = 0)
survdiff(Surv(Time,status) ~ media, data = d_fly, rho = 1)
fit_fly <- coxph(Surv(Time,status) ~media, data = d_fly)
summary(fit_fly)
```

Then 20 degrees and OP50
```{r}
survdiff(Surv(Time,status) ~ media, data = d_2050, rho = 0)
survdiff(Surv(Time,status) ~ media, data = d_2050, rho = 1)
fit_2050 <- coxph(Surv(Time,status) ~media, data = d_2050)
summary(fit_2050)
```

Then for 13 degrees and OP50
```{r}
survdiff(Surv(Time,status) ~ media, data = d_1350, rho = 0)
survdiff(Surv(Time,status) ~ media, data = d_1350, rho = 1)
fit_1350 <- coxph(Surv(Time,status) ~media, data = d_1350)
summary(fit_1350)
```

 Where only 20 degrees and 13 degrees are significantly different from each other. But there is also some difference between 20 degrees and OP50. Where there isn´t a big difference between 13 degrees and
 OP50.
 
 Then dose response curves with first the two fly media.
 
```{r}
plot(df_fly_13$Time[2:9],df_fly_13$Surv[2:9], pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival heat stress for'~italic(C.elegans)),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
lines(df_fly_13$Time,df_fly_13$Surv)
arrows(df_fly_13$Time, df_fly_13$Surv-df_fly_13$Std.error, df_fly_13$Time, 
       df_fly_13$Surv+df_fly_13$Std.error, length=0.05, angle=90, code=3, col = 'black')
points(df_fly_20$Time[2:9],df_fly_20$Surv[2:9], pch = 6)
lines(df_fly_20$Time,df_fly_20$Surv,lty = 3)
arrows(df_fly_20$Time, df_fly_20$Surv-df_fly_20$Std.error, df_fly_20$Time, 
       df_fly_20$Surv+df_fly_20$Std.error, length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,6),lty = c(1,3), 
       legend = c(expression('Fly gut 13'*degree*C),
                  expression('Fly gut 20'*degree*C)), bty = 'n')
```

Than 20 degrees and OP50

```{r}
plot(df_fly_20$Time[2:9],df_fly_20$Surv[2:9], pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival heat stress for'~italic(C.elegans)),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
points(df_OP50$Time[2:9],df_OP50$Surv[2:9], pch = 1, cex = 1.2)
lines(df_OP50$Time,df_OP50$Surv, lty = 2)
arrows(df_OP50$Time, df_OP50$Surv-df_OP50$Std.error,
      df_OP50$Time, df_OP50$Surv+df_OP50$Std.error, 
      length=0.05, angle=90, code=3, col = 'black')
lines(df_fly_20$Time,df_fly_20$Surv,lty = 1)
arrows(df_fly_20$Time, df_fly_20$Surv-df_fly_20$Std.error, df_fly_20$Time, 
       df_fly_20$Surv+df_fly_20$Std.error, length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,1),lty = c(1,2), 
       legend = c(expression('Fly gut 20'*degree*C),
                  expression(italic(E.coli) ~ OP50)), bty = 'n')
```

Then 13 degrees and OP50
```{r}
plot(df_fly_13$Time[2:9],df_fly_13$Surv[2:9], pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival heat stress for'~italic(C.elegans)),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
lines(df_fly_13$Time,df_fly_13$Surv)
arrows(df_fly_13$Time, df_fly_13$Surv-df_fly_13$Std.error, df_fly_13$Time, 
       df_fly_13$Surv+df_fly_13$Std.error, length=0.05, angle=90, code=3, col = 'black')
points(df_OP50$Time[2:9],df_OP50$Surv[2:9], pch = 1, cex = 1.2)
lines(df_OP50$Time,df_OP50$Surv, lty = 2)
arrows(df_OP50$Time, df_OP50$Surv-df_OP50$Std.error,
      df_OP50$Time, df_OP50$Surv+df_OP50$Std.error, 
      length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,1),lty = c(1,2), 
       legend = c(expression('Fly gut 13'*degree*C),
                  expression(italic(E.coli) ~ OP50)), bty = 'n')
```
Then Kaplan meier curves
```{r}
km_fly <- survfit(Surv(Time,status) ~ media, data = d_fly,conf.type = "log-log", error = "greenwood")
km_2050 <- survfit(Surv(Time,status) ~ media, data = d_2050,conf.type = "log-log", error = "greenwood")
km_1350 <- survfit(Surv(Time,status) ~ media, data = d_1350,conf.type = "log-log", error = "greenwood")
ggsurvplot(km_fly,data = d_fly, conf.int = TRUE,
           ggtheme = theme_bw(),risk.table = 0.25,
           palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media',break.x.by = 2)
ggsurvplot(km_2050,data = d_2050, conf.int = TRUE, 
          ggtheme = theme_bw(), risk.table = 0.25,
          palette ='aaas', surv.median.line = 'hv',
          legend.title = 'Media',break.x.by = 2)
ggsurvplot(km_1350,data = d_1350, conf.int = TRUE, 
           ggtheme = theme_bw(),risk.table = 0.25,
           palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media', break.x.by = 2) 
```

 
We then try to split up in replicates to see if there are any differences
between scorers

```{r}
d1 <- data[data$Replicate == 1,]
d2 <- data[data$Replicate == 2,]
d1$Survobj <- with(d1,Surv(d1$Time, event = d1$status))
d2$Survobj <- with(d2,Surv(d2$Time, event = d2$status))
```

Then we make kaplan-meier curves and dose response curves

```{r}

km1 <- survfit(Survobj ~ media, data = d1,conf.type = "log-log")
s_km1 <- summary(km1)
km2 <- survfit(Survobj ~ media, data = d2,conf.type = "log-log")
s_km2 <- summary(km2)
df_fly_13 <- data.frame(c(0,s_km1$time[1:7]),c(1,s_km1$surv[1:7]),
                        c(0,s_km1$std.err[1:7]))
colnames(df_fly_13) <- c('Time','Surv','Std.error')
df_fly_20 <- data.frame(c(0,s_km1$time[8:15]),c(1,s_km1$surv[8:15]),
                        c(0,s_km1$std.err[8:15]))
colnames(df_fly_20) <- c('Time','Surv','Std.error')
df_OP50 <- data.frame(c(0,s_km1$time[16:24]),c(1,s_km1$surv[16:24]),
                        c(0,s_km1$std.err[16:24]))
colnames(df_OP50) <- c('Time','Surv','Std.error')

plot(df_fly_13$Time[2:9],df_fly_13$Surv[2:9], pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival heat stress for'~italic(C.elegans)),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
lines(df_fly_13$Time,df_fly_13$Surv)
arrows(df_fly_13$Time, df_fly_13$Surv-df_fly_13$Std.error, df_fly_13$Time, 
       df_fly_13$Surv+df_fly_13$Std.error, length=0.05, angle=90, code=3, col = 'black')
points(df_OP50$Time[2:9],df_OP50$Surv[2:9], pch = 1, cex = 1.2)
lines(df_OP50$Time,df_OP50$Surv, lty = 2)
arrows(df_OP50$Time, df_OP50$Surv-df_OP50$Std.error,
      df_OP50$Time, df_OP50$Surv+df_OP50$Std.error, 
      length=0.05, angle=90, code=3, col = 'black')
points(df_fly_20$Time[2:9],df_fly_20$Surv[2:9], pch = 6)
lines(df_fly_20$Time,df_fly_20$Surv,lty = 3)
arrows(df_fly_20$Time, df_fly_20$Surv-df_fly_20$Std.error, df_fly_20$Time, 
       df_fly_20$Surv+df_fly_20$Std.error, length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,6,1),lty = c(1,3,2), 
       legend = c(expression('Fly gut 13'*degree*C),
                  expression('Fly gut 20'*degree*C),
                  expression(italic(E.coli) ~ OP50)), bty = 'n')

df_fly_13 <- data.frame(c(0,s_km2$time[1:6],18),
                        c(1,s_km2$surv[1:6],s_km2$surv[6]),
                        c(0,s_km2$std.err[1:6],s_km2$std.err[6]))
colnames(df_fly_13) <- c('Time','Surv','Std.error')
df_fly_20 <- data.frame(c(0,s_km2$time[7:11]),c(1,s_km2$surv[7:11]),
                        c(0,s_km2$std.err[7:11]))
colnames(df_fly_20) <- c('Time','Surv','Std.error')
df_OP50 <- data.frame(c(0,s_km2$time[12:19]),c(1,s_km2$surv[12:19]),
                        c(0,s_km2$std.err[12:19]))
colnames(df_OP50) <- c('Time','Surv','Std.error')

plot(df_fly_13$Time[2:8],df_fly_13$Surv[2:8],
     pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival heat stress for'~italic(C.elegans)),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
lines(df_fly_13$Time,df_fly_13$Surv)
arrows(df_fly_13$Time, df_fly_13$Surv-df_fly_13$Std.error, df_fly_13$Time, 
       df_fly_13$Surv+df_fly_13$Std.error, length=0.05, angle=90, code=3, col = 'black')
points(df_OP50$Time[2:9],df_OP50$Surv[2:9], pch = 1, cex = 1.2)
lines(df_OP50$Time,df_OP50$Surv, lty = 2)
arrows(df_OP50$Time, df_OP50$Surv-df_OP50$Std.error,
      df_OP50$Time, df_OP50$Surv+df_OP50$Std.error, 
      length=0.05, angle=90, code=3, col = 'black')
points(df_fly_20$Time[2:9],df_fly_20$Surv[2:9], pch = 6)
lines(df_fly_20$Time,df_fly_20$Surv,lty = 3)
arrows(df_fly_20$Time, df_fly_20$Surv-df_fly_20$Std.error, df_fly_20$Time, 
       df_fly_20$Surv+df_fly_20$Std.error, length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,6,1),lty = c(1,3,2), 
       legend = c(expression('Fly gut 13'*degree*C),
                  expression('Fly gut 20'*degree*C),
                  expression(italic(E.coli) ~ OP50)), bty = 'n')
```

Then kaplan meier
```{r}
ggsurvplot(km1,data = data, conf.int = TRUE, ggtheme = theme_bw(),
           risk.table = 0.25, palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media',break.x.by = 2) 
ggsurvplot(km2,data = data, conf.int = TRUE, ggtheme = theme_bw(),
           risk.table = 0.25, palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media',break.x.by = 2) 
```
 and at last some tests
 
```{r}
survdiff(Survobj ~media, data = d1, rho = 0) 
survdiff(Survobj ~media, data = d1, rho = 1) 
fit_d1 <- coxph(Survobj ~media, data = d1)
summary(fit_d1)
survdiff(Survobj ~media, data = d2, rho = 0) 
survdiff(Survobj ~media, data = d2, rho = 1) 
fit_d2 <- coxph(Survobj ~media, data = d2)
summary(fit_d2)
```

Now we use the replicate dataset
```{r}
data_2 <- read.table(here('data', 'worm_second.csv'), header = TRUE,
                     sep = ';')
head(data_2)
colnames(data_2) <- c('Time','media','Metabolite','Status')
data_2_rep <- subset(data_2, Metabolite == 0)
```

First we try to see if we can extend the cox model from the earlier dataset to the new one to see hoow well we can extrapolate from one dataset to  the other

```{r}
plot(survfit(fit, newdata = data.frame(media = 'FG20')), lty = 2, xaxp = c(0,18,9), col = 'blue',
     main = 'Predictions for model and sample data',
     xlab = 'hours heat stress')
lines(survfit(fit, newdata = data.frame(media = 'FG13')), col = 'firebrick')
lines(survfit(fit, newdata = data.frame(media = 'OP50')), lty =3)
points(df_OP50$Time, df_OP50$Surv, col = 'black', pch = 16)
points(df_fly_13$Time,df_fly_13$Surv, col = 'firebrick', pch = 16)
points(df_fly_20$Time,df_fly_20$Surv, col = 'blue', pch = 16)
legend('topright', pch = 16,lty = c(1,3,2), 
       col = c('firebrick','blue','black'), 
       legend = c(expression('Fly gut 13'*degree*C),
                  expression('Fly gut 20'*degree*C),
                  expression(italic(E.coli) ~ OP50)), bty = 'n')
```

Here the data fits well if we do the same for the new data set.

```{r}
data_2_rep$Survobj <- with(data_2_rep,
                           Surv(data_2_rep$Time, event = data_2_rep$Status))
km_2 <- survfit(Survobj ~ media, data = data_2_rep,conf.type = "log-log", 
              error = "greenwood")
s_km_2 <- summary(km_2)
ggsurvplot(km_2)
fit_2 <- coxph(Survobj ~ media, data = data_2_rep)
summary(fit_2)
```
We see that Fly gut 20 has extremely bad survival and our former predictions do not hold for the new dataset.

If we try to plot the new data for old predictions

```{r}
df_fly_13_2 <- data.frame(c(0,s_km_2$time[1:8]),c(1,s_km_2$surv[1:8]),
                        c(0,s_km_2$std.err[1:8]))
colnames(df_fly_13_2) <- c('Time','Surv','Std.error')
df_fly_20_2 <- data.frame(c(0,s_km_2$time[9:14]),c(1,s_km_2$surv[9:14]),
                        c(0,s_km_2$std.err[9:14]))
colnames(df_fly_20_2) <- c('Time','Surv','Std.error')
df_OP50_2 <- data.frame(c(0,s_km_2$time[15:21]),c(1,s_km_2$surv[15:21]),
                        c(0,s_km_2$std.err[15:21]))
colnames(df_OP50_2) <- c('Time','Surv','Std.error')
```

```{r}
plot(survfit(fit, newdata = data.frame(media = 'FG20')), lty = 2, xaxp = c(0,18,9), col = 'blue',
     main = 'Predictions from old model and new data points',
     xlab = 'hours heat stress')
lines(survfit(fit, newdata = data.frame(media = 'FG13')), col = 'firebrick')
lines(survfit(fit, newdata = data.frame(media = 'OP50')), lty =3)
points(df_OP50_2$Time, df_OP50_2$Surv, col = 'black', pch = 16)
points(df_fly_13_2$Time,df_fly_13_2$Surv, col = 'firebrick', pch = 16)
points(df_fly_20_2$Time,df_fly_20_2$Surv, col = 'blue', pch = 16)
legend('topright', pch = 16,lty = c(1,3,2), 
       col = c('firebrick','blue','black'),
       legend = c(expression('Fly gut 13'*degree*C),
                  expression('Fly gut 20'*degree*C),
                  expression(italic(E.coli) ~ OP50)), bty = 'n')
```
 Both FG13 and OP50 is slightly lower than before but the survival in F20 has cratered and this may have happened because of something which liquefied the 
 worms on the plates.

Then we compare data with standard errors with one group in each plot 
```{r}
plot(df_fly_13$Time[2:9],df_fly_13$Surv[2:9], pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival for' ~ italic(C.elegans) ~' from flygut 13' ~degree*C),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
lines(df_fly_13$Time,df_fly_13$Surv)
arrows(df_fly_13$Time, df_fly_13$Surv-df_fly_13$Std.error, df_fly_13$Time, 
       df_fly_13$Surv+df_fly_13$Std.error, length=0.05, angle=90, code=3, col = 'black')
points(df_fly_13_2$Time[2:9],df_fly_13_2$Surv[2:9], pch = 1, cex = 1.2) 
lines(df_fly_13_2$Time,df_fly_13_2$Surv, lty = 2)
arrows(df_fly_13_2$Time, df_fly_13_2$Surv-df_fly_13_2$Std.error, df_fly_13_2$Time, df_fly_13_2$Surv+df_fly_13_2$Std.error, length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,1),lty = c(1,2), 
       legend = c('Old dataset', 'New dataset'), bty = 'n')
```
 
```{r}
plot(df_fly_20$Time[2:9],df_fly_20$Surv[2:9], pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival for' ~ italic(C.elegans) ~' from flygut 20' ~degree*C),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
lines(df_fly_20$Time,df_fly_20$Surv)
arrows(df_fly_20$Time, df_fly_20$Surv-df_fly_20$Std.error, df_fly_20$Time, 
       df_fly_20$Surv+df_fly_20$Std.error, length=0.05, angle=90, code=3, col = 'black')
points(df_fly_20_2$Time[2:9],df_fly_20_2$Surv[2:9], pch = 1, cex = 1.2) 
lines(df_fly_20_2$Time,df_fly_20_2$Surv, lty = 2)
arrows(df_fly_20_2$Time, df_fly_20_2$Surv-df_fly_20_2$Std.error, df_fly_20_2$Time, df_fly_20_2$Surv+df_fly_20_2$Std.error, length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,1),lty = c(1,2), 
       legend = c('Old dataset', 'New dataset'), bty = 'n')
```
 
```{r}
plot(df_OP50$Time[2:9],df_OP50$Surv[2:9], pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival for' ~ italic(C.elegans) ~ 'from' ~ italic(E.coli) ~ OP50),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
lines(df_OP50$Time,df_OP50$Surv)
arrows(df_OP50$Time, df_OP50$Surv-df_OP50$Std.error, df_OP50$Time, 
       df_OP50$Surv+df_OP50$Std.error, length=0.05, angle=90, code=3, col = 'black')
points(df_OP50_2$Time[2:9],df_OP50_2$Surv[2:9], pch = 1, cex = 1.2) 
lines(df_OP50_2$Time,df_OP50_2$Surv, lty = 2)
arrows(df_OP50_2$Time, df_OP50_2$Surv-df_OP50_2$Std.error, df_OP50_2$Time, df_OP50_2$Surv+df_OP50_2$Std.error, length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,1),lty = c(1,2), 
       legend = c('Old dataset', 'New dataset'), bty = 'n')
```

