---
title: "Survival of C.elegans with three different media"
author: "Jonas Gehrlein"
date: "12 nov 2018"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(survival) 
library(here)
library(survminer)
```
code for fitting Kaplan-Meier and log-rank test and for displaying 
survival curves for each type of media

```{r}
data <- read.table(here('data','Worm_4days.csv'), header = TRUE, sep = ";", dec = ",")
head(data)
```

here() should show that your position is in the folder 7.semester else select the active project to be 7.semester in the upper right corner. 

Or make a new .Rproj file in the folder 7.semester. There is a problem with the numbers of the worms so we change that with col.names()
```{r}
colnames(data)<-c("Time","status","media",'Replicate')
head(data)
```

Then we create a survival object with the survival package and make a kaplan-meier curve
```{r}
data$Survobj <- with(data,Surv(data$Time, event = data$status))
km <- survfit(Survobj ~ media, data = data,conf.type = "log-log", error = "greenwood")
s_km <- summary(km)
plot(km, col = c("red","black",'blue'))
```

Shows errorbars and lineplot based on one of Anders papers https://onlinelibrary.wiley.com/doi/full/10.1111/acel.12165
This is done by extracting the surviving proportion, standard errors and time from survfit().  and adding a startpoint where the survival is 100% at 0 hours.

```{r}
df_fly_13 <- data.frame(c(0,s_km$time[1:8]),c(1,s_km$surv[1:8]),
                        c(0,s_km$std.err[1:8]))
colnames(df_fly_13) <- c('Time','Surv','Std.error')
df_fly_20 <- data.frame(c(0,s_km$time[9:16]),c(1,s_km$surv[9:16]),
                        c(0,s_km$std.err[9:16]))
colnames(df_fly_20) <- c('Time','Surv','Std.error')
df_OP50 <- data.frame(c(0,s_km$time[17:24]),c(1,s_km$surv[17:24]),
                        c(0,s_km$std.err[17:24]))
colnames(df_OP50) <- c('Time','Surv','Std.error')

plot(df_fly_13$Time[2:9],df_fly_13$Surv[2:9], pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival heat stress for'~italic(C.elegans)),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
lines(df_fly_13$Time,df_fly_13$Surv)
arrows(df_fly_13$Time, df_fly_13$Surv-df_fly_13$Std.error, df_fly_13$Time, 
       df_fly_13$Surv+df_fly_13$Std.error, length=0.05, angle=90, code=3, col = 'black')
points(df_OP50$Time[2:9],df_OP50$Surv[2:9], pch = 1, cex = 1.2)
lines(df_OP50$Time,df_OP50$Surv, lty = 2)
arrows(df_OP50$Time, df_OP50$Surv-df_OP50$Std.error,
      df_OP50$Time, df_OP50$Surv+df_OP50$Std.error, 
      length=0.05, angle=90, code=3, col = 'black')
points(df_fly_20$Time[2:9],df_fly_20$Surv[2:9], pch = 6)
lines(df_fly_20$Time,df_fly_20$Surv,lty = 3)
arrows(df_fly_20$Time, df_fly_20$Surv-df_fly_20$Std.error, df_fly_20$Time, 
       df_fly_20$Surv+df_fly_20$Std.error, length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,6,1),lty = c(1,3,2), 
       legend = c(expression('Fly gut 13'*degree*C),
                  expression('Fly gut 20'*degree*C),
                  expression(italic(E.coli) ~ OP50)), bty = 'n')
```

Then we create a normal Kaplan-Meier curve
```{r}
ggsurvplot(km,data = data, conf.int = TRUE, ggtheme = theme_bw(),
           risk.table = 0.25, palette ='aaas', surv.median.line = 'hv',legend.title = 'Media',break.x.by = 2) 
```

Now we test for difference between the curves with both log-rank and gehan-wilcoxon and with an cox proportional hazard model

```{r}
survdiff(Survobj ~media, data = data, rho = 0) 
survdiff(Survobj ~media, data = data, rho = 1) 
fit <- coxph(Survobj ~media, data = data)
summary(fit)
```




We then try to split up in replicates to see if there are any differences
between scorers

```{r}
d1 <- data[data$Replicate == 1,]
d2 <- data[data$Replicate == 2,]
d1$Survobj <- with(d1,Surv(d1$Time, event = d1$status))
d2$Survobj <- with(d2,Surv(d2$Time, event = d2$status))
```

Then we make kaplan-meier curves and dose response curves

```{r}

km1 <- survfit(Survobj ~ media, data = d1,conf.type = "log-log")
s_km1 <- summary(km1)
km2 <- survfit(Survobj ~ media, data = d2,conf.type = "log-log")
s_km2 <- summary(km2)
df_fly_13 <- data.frame(c(0,s_km1$time[1:7]),c(1,s_km1$surv[1:7]),
                        c(0,s_km1$std.err[1:7]))
colnames(df_fly_13) <- c('Time','Surv','Std.error')
df_fly_20 <- data.frame(c(0,s_km1$time[8:15]),c(1,s_km1$surv[8:15]),
                        c(0,s_km1$std.err[8:15]))
colnames(df_fly_20) <- c('Time','Surv','Std.error')
df_OP50 <- data.frame(c(0,s_km1$time[16:24]),c(1,s_km1$surv[16:24]),
                        c(0,s_km1$std.err[16:24]))
colnames(df_OP50) <- c('Time','Surv','Std.error')

plot(df_fly_13$Time[2:9],df_fly_13$Surv[2:9], pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival heat stress for'~italic(C.elegans)),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
lines(df_fly_13$Time,df_fly_13$Surv)
arrows(df_fly_13$Time, df_fly_13$Surv-df_fly_13$Std.error, df_fly_13$Time, 
       df_fly_13$Surv+df_fly_13$Std.error, length=0.05, angle=90, code=3, col = 'black')
points(df_OP50$Time[2:9],df_OP50$Surv[2:9], pch = 1, cex = 1.2)
lines(df_OP50$Time,df_OP50$Surv, lty = 2)
arrows(df_OP50$Time, df_OP50$Surv-df_OP50$Std.error,
      df_OP50$Time, df_OP50$Surv+df_OP50$Std.error, 
      length=0.05, angle=90, code=3, col = 'black')
points(df_fly_20$Time[2:9],df_fly_20$Surv[2:9], pch = 6)
lines(df_fly_20$Time,df_fly_20$Surv,lty = 3)
arrows(df_fly_20$Time, df_fly_20$Surv-df_fly_20$Std.error, df_fly_20$Time, 
       df_fly_20$Surv+df_fly_20$Std.error, length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,6,1),lty = c(1,3,2), 
       legend = c(expression('Fly gut 13'*degree*C),
                  expression('Fly gut 20'*degree*C),
                  expression(italic(E.coli) ~ OP50)), bty = 'n')

df_fly_13 <- data.frame(c(0,s_km2$time[1:6],18),
                        c(1,s_km2$surv[1:6],s_km2$surv[6]),
                        c(0,s_km2$std.err[1:6],s_km2$std.err[6]))
colnames(df_fly_13) <- c('Time','Surv','Std.error')
df_fly_20 <- data.frame(c(0,s_km2$time[7:11]),c(1,s_km2$surv[7:11]),
                        c(0,s_km2$std.err[7:11]))
colnames(df_fly_20) <- c('Time','Surv','Std.error')
df_OP50 <- data.frame(c(0,s_km2$time[12:19]),c(1,s_km2$surv[12:19]),
                        c(0,s_km2$std.err[12:19]))
colnames(df_OP50) <- c('Time','Surv','Std.error')

plot(df_fly_13$Time[2:8],df_fly_13$Surv[2:8],
     pch = 16, cex = 1.2, xlab = 'Heatstress (hours)', 
     ylab = 'Surviving fraction', 
     main = expression('Survival heat stress for'~italic(C.elegans)),
     xlim = c(0,22), ylim = c(0,1), xaxp = c(0,22,11))
lines(df_fly_13$Time,df_fly_13$Surv)
arrows(df_fly_13$Time, df_fly_13$Surv-df_fly_13$Std.error, df_fly_13$Time, 
       df_fly_13$Surv+df_fly_13$Std.error, length=0.05, angle=90, code=3, col = 'black')
points(df_OP50$Time[2:9],df_OP50$Surv[2:9], pch = 1, cex = 1.2)
lines(df_OP50$Time,df_OP50$Surv, lty = 2)
arrows(df_OP50$Time, df_OP50$Surv-df_OP50$Std.error,
      df_OP50$Time, df_OP50$Surv+df_OP50$Std.error, 
      length=0.05, angle=90, code=3, col = 'black')
points(df_fly_20$Time[2:9],df_fly_20$Surv[2:9], pch = 6)
lines(df_fly_20$Time,df_fly_20$Surv,lty = 3)
arrows(df_fly_20$Time, df_fly_20$Surv-df_fly_20$Std.error, df_fly_20$Time, 
       df_fly_20$Surv+df_fly_20$Std.error, length=0.05, angle=90, code=3, col = 'black')
legend('topright', pch = c(16,6,1),lty = c(1,3,2), 
       legend = c(expression('Fly gut 13'*degree*C),
                  expression('Fly gut 20'*degree*C),
                  expression(italic(E.coli) ~ OP50)), bty = 'n')
```

Then kaplan meier
```{r}
ggsurvplot(km1,data = data, conf.int = TRUE, ggtheme = theme_bw(),
           risk.table = 0.25, palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media',break.x.by = 2) 
ggsurvplot(km2,data = data, conf.int = TRUE, ggtheme = theme_bw(),
           risk.table = 0.25, palette ='aaas', surv.median.line = 'hv',
           legend.title = 'Media',break.x.by = 2) 
```
 and at last some tests
 
```{r}
survdiff(Survobj ~media, data = d1, rho = 0) 
survdiff(Survobj ~media, data = d1, rho = 1) 
fit <- coxph(Survobj ~media, data = d1)
summary(fit)
survdiff(Survobj ~media, data = d2, rho = 0) 
survdiff(Survobj ~media, data = d2, rho = 1) 
fit <- coxph(Survobj ~media, data = d2)
summary(fit)
```
 
